---
title: "CompoundDb"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

https://rformassspectrometry.github.io/CompoundDb/articles/create-compounddb.html#compdb-from-custom-data

```{r, include = FALSE}
startpoint <- Sys.time()
```


# Libraries

```{r, message=FALSE}
db <- "AnnComp" # "stds" "SL" "MS2_ann", AnnComp

library(readxl)
library(MetaboCoreUtils)
library(CompoundDb)
```


# Data import

```{r import}
if(file.exists(paste0("rdata/CompDb_", db, ".sqlite"))){
  file.remove(paste0("rdata/CompDb_", db, ".sqlite"))
}
if(db == "stds"){
  cmps <- read_xlsx("data/standards.xlsx")
} else if (db == "SL"){
  cmps <- read.csv("data/SuspectList_3rd_200223.csv")
} else if (db == "AnnComp"){
  cmps <- read_xlsx("data/AnnotatedComp_FINAL_ionDB.xlsx")
} else if(db == "MS2_ann"){
  cmps <- read_xlsx("data/AnnotatedCompounds_SL.xlsx")
}

cmps$mass <- calculateMass(cmps$formula)
```


# Compound DB building

Create a `CompDb` database from automatically defined compounds.  
Create a `data.frame` with compound information of candidate neuroactive compound.   
The data needs to have the following information "compound_id", "name", 
"inchi", "inchikey", "formula", "exactmass", "synonyms".  
All columns except "compound_id" can also contain missing values.    
It is also possible to define additional columns.  

```{r cmpdb}
if(db == "SL"){
  tmp_rt <- rep(0, nrow(cmps))
} else {
  tmp_rt <- cmps$RT
}

cmps <- data.frame(
  compound_id = paste0(
    "C", formatC(seq(nrow(cmps)), width = nchar(nrow(cmps)), flag = "0")),
  name = cmps$name,
  inchi = NA_character_,
  inchikey = NA_character_,
  formula = cmps$formula,
  exactmass = cmps$mass,
  synonyms = NA_character_,
  rtime = tmp_rt,
  adduct = c("[M-H]-; [M+H]+"),
  comment = cmps$comment,
  activity = cmps$activity,
 level = cmps$level,
 mode = cmps$mode,
 abc = cmps$group
)
```


It is also required to define *metadata* for the database with the 
`make_metadata` function.

```{r metad}
metad <- make_metadata(source = "pomodoro", url = "",
                       source_version = "", 
                       source_date = Sys.Date(),
                       organism = NA_character_)
```

With this information we can already create the `CompDb` database.

```{r dbfile}
db_file <- createCompDb(cmps, metadata = metad, path = "rdata/", 
                        dbFile = paste0("CompDb_", db, ".sqlite"))
```

Below we load this database using the `CompDb` function in read-write mode 
because we want to add more information to this database later.

```{r cdb}
(cdb <- CompDb(db_file, flags = RSQLite::SQLITE_RW))
```

Next step is to create and `IonDb` database, which extends the `CompDb` by 
adding additional database tables to support ion annotations.

```{r idb}
idb <- IonDb(cdb)
```

Below we define a data frame with ion annotations and add that to the database 
with the `insertIon` function. The ion annotations are associated with the 
class of lipid. The column "compound_id" needs to contain the identifiers of 
the compounds to which the ion should be related to.

```{r idb2}
for(i in seq(nrow(cmps))){
  x <- unlist(strsplit(cmps$adduct[i], "; "))
  ion <- data.frame(
    compound_id = rep(cmps$compound_id[i], length(x)),
    ion_adduct = x,
    ion_mz = rep(NA, length(x)),
    ion_rt = rep(cmps$rtime[i], length(x))
  )
  for(j in seq(nrow(ion))){
    ion$ion_mz[j] <- mass2mz(cmps$exactmass[i], x[j])
  }
  idb <- insertIon(idb, ion)
}
```


# Save the output

```{r save}
ionsdb <- ions(idb, c(
  "compound_id", "name", "ion_adduct", "ion_mz", "ion_rt", "comment", "activity", "level", "mode", "abc"))
save(ionsdb, file = paste0("rdata/ionsdb_", db, ".RData"))
```


# Session info

```{r session-info}
Sys.time()-startpoint
devtools::session_info()
```

