---
title: "Data workflow"
author: "Mar Garcia-Aloy"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: false
    number_sections: false
    toc_float: false
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In this document we do the workflow of feature selection to identify interesting compounds.  
From original dataset from Progenesis and processed by AK we select the good quality features.   
Then we match their mz values with the mz of suspect list (considering only [M+H]+) and then we keep those features that matched with at least 1 suspect mz value.  
  
**Result:** 
  POSITIVE MODE
  - from 20.915 features --> 17.207 good quality features  
  - from 17.207 features --> 2.825 m/z matching with suspect list   
  - from  2.825 features --> 58 annotated compounds   

```{r}
library(tidyverse)
library(MetaboCoreUtils)
library(MetaboAnnotation)
library(effectsize)

# Suspect list import and m/z calculation:
slist <- read.csv("data/SuspectList_3rd_200223.csv")
slist$mass <- calculateMass(slist$formula)
idx <- which(is.na(slist$mass))
if(length(idx) > 0){
  slist <- slist[-idx,]
}
slist$mz <- as.numeric(mass2mz(slist$mass, "[M+H]+"))


# Study data import (processed by AK):
load("data/mydat_neg.RData")
nrow(mydat)
table(mydat$noise) # from 20.915 features --> 17.207 good quality features

# Match study features with suspect list m/z:
feats <- mydat %>%
  filter(noise == FALSE) %>%
  select(name, mz)
mymatchs <- matchValues(data.frame(feats),
                        data.frame(slist[,c("name", "mz", "activity")]), 
                        param = MzParam(tolerance = 0.01))
mymatchdt <- data.frame(matchedData(mymatchs))

mymatchdt <- mymatchdt %>%
  as_tibble() %>%
  filter(!is.na(target_name)) %>%
  nest(suspect = !c(name, mz))

mymatchdt <- mymatchdt %>%
  mutate(suspect_n = map_dbl(suspect, function(t) nrow(t)))

mydat <- mydat %>%
  left_join(mymatchdt %>% select(!mz))
rm(slist, mymatchs, mymatchdt, feats, idx)

mydat <- mydat %>%
  filter(noise == FALSE & !is.na(suspect_n))
nrow(mydat) # from 17.207 good quality features --> 2.825 m/z matching with suspect list

# Get maximum mean(group):
mydat <- mydat %>%
  mutate(data = map(data, function(t){
    t %>%
      mutate(norm_indiv_0 = ifelse(is.na(norm_indiv), 0, norm_indiv)) %>%
      separate(id, sep = "_", into = c("order", "sample", "rep1", "rep2"), remove = FALSE) %>%
      mutate(group = gsub(paste(seq(6), collapse = "|"), "", sample))
  })) %>%
  mutate(mean_group = map(data, function(t){
    t %>% 
      group_by(group) %>% 
      summarise(mean_log  = mean(norm_indiv_i))
  })) %>%
  mutate(mean_max = map_dbl(mean_group, function(t){
    t %>%
      filter(group %in% c("D", "DO", "PF", "PS")) %>%
      pull(mean_log) %>%
      max(.)
  }))
```

# Session information
```{r session}
Sys.time()-startpoint
devtools::session_info()
```