---
title: "Data Analysis"
author: "Ana Kovacic"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
```


# Libraries

```{r}
p <- "POS"
software <- "progenesis" #progenesis / mzmine
myfile <- "data/ANA_MSE_POS_ALL_MASS_PROG.xlsx" #"data/ANA_MSE_NEG_ALL_MZMINE.xlsx" #"data/ANA_MSE_POS_ALL_MASS_PROG.xlsx", "data/ANA_MSE_NEG_ALL_MASS_PROG.xlsx" 

library(writexl)
library(readxl)
library(tidyverse)
library(plotly)
library(MetaboAnnotation)
library(effectsize)
library(DT)
library(kableExtra)
library(vegan)
library(devtools)
#library(ggvegan)
library(reshape2)
library(officer) 
library(VennDiagram)
library(grid)
library(officer) 


scaling.pareto <- function(x){(x - mean(x)) / sqrt(sd(x))}
```


# Data import

```{r, message=FALSE}
if(software == "progenesis"){
  feat <- read_excel(myfile, skip = 2) #SKIP first two empty rows 
  
  tmp <- read_excel(myfile)
  
  # select raw data:
  data_raw <- feat[, grep("Raw", colnames(tmp)):(grep("Not identified", tmp[2,])-1)]
  
  feat <- data.frame(feat[,c(3, 5)])
  colnames(feat) <- c("mz", "RT")
  
  rm(tmp)
} else if(software == "mzmine"){
  feat <- read_xlsx(myfile) 
  
  data_raw <- feat[,4:ncol(feat)]
  feat <- data.frame(feat[,c(2, 3)])
  colnames(feat) <- c("mz", "RT")
}
feat$FT <- paste0(
  "FT", formatC(seq(nrow(feat)), width = nchar(nrow(feat)), flag = "0"))
rownames(data_raw) <- feat$FT
data_raw <- data.frame(t(data_raw))
```

# Data organization

```{r}
smpl_class <- rep("sample", nrow(data_raw))
smpl_class[grep("QC", rownames(data_raw))] <- "QC"
smpl_class[grep("Blank|PPB", rownames(data_raw))] <- "blank"

# number of NON-missing values in QC
idx_QC <- grep("QC", rownames(data_raw))
smpl_order <- as.numeric(substr(rownames(data_raw), 13, 14))
idx_QC <- idx_QC[(smpl_order[idx_QC] >= 8) & (smpl_order[idx_QC] <= 59)] #take only qc from the last of the first and the first of the last
data_QC <- data_raw[idx_QC, ]
feat$QC_nonNA <- apply(data_QC, 2, function(x) sum(x > 0)) #in how many samples the feature is not zero, missing?

rm(idx, idx_QC, data_QC, smpl_order) 

metadata <- data.frame(
  id = gsub(".*_AK_|_MSe.*", "", rownames(data_raw)) #id is the sample name
) #we are using tidevrsy because the variables can be different and not only numeric like should be in dataframe
metadata$order <- as.numeric(gsub("_.*", "", metadata$id)) # second colum order of samples
metadata$class <- smpl_class #third colume is sample class

mydat <- cbind.data.frame(metadata, data_raw) #we have bind now the metadata with raw data

colnames(feat)[colnames(feat) == "FT"] <- "name"
mydat <- mydat %>%
  as.tibble() %>% # transform the object to a "tibble" 
  # each row of the tibble corresponds to a compound:
  pivot_longer(starts_with("FT"))  %>% 
  # put all columns, except "name", in the list of the tibble called "data"
  nest(data = !c("name")) %>% #this information will be in gather in one column - called data
  # aggregate feature information
  left_join(feat, by = "name") #information on id, order, class,value - raw
rm(data_raw, feat, metadata, smpl_class) # remove the non needing things 

mydat <- mydat %>%
  mutate(data = map(data, function(t){
    t$val_0 <- t$value
      t$value <- as.numeric(t$value) #in case of mzmine data
    t$value[t$value == 0] <- NA #replace missing values with zero
    t$log <- log10(t$value) #transform the t_value, log10 of zero return NA
    t
  }))

#mydat <- mydat[!mydat$noise, ] #for now we decide not to exclude the noise, based on preliminary result
```


# Normalize

```{r}
mydat <- mydat %>% #we decide not to impute data before normalisation 
  mutate(norm_mdl = map(data, function(t){
    t <- t %>%
      filter(class == "QC" & order >= 8 & order <= 59) #for normalisation we use the last of the first qc in between and the one before last one
    if(sum(is.na(t$log)) < nrow(t)*3/4){ #if feature not missing in more than 75% samples than apply the function?
      val <- t$log #we transform data before normalisation
      order <- t$order
      mdl <- lm(val ~ order) #we apply the linear normalisation on the values withing the order of injection
      mdl
    }
  }))

myimputer <- function(v, w){
  set.seed(123) #change the missing values with the random value between min value of the colume/10
  if(sum(is.na(v)) == 0){
    return(v)
  } else {
    napos <- which(is.na(v))
    newval <- runif(length(napos), 0, min(w, na.rm = TRUE)/10)
    out <- v
    out[napos] <- log10(newval) #tranform the new data
    return(out)
  }
}
mydat <- mydat %>% #normalise by using the individual slope exluding the missing values
  mutate(data = map2(data, norm_mdl, function(t, u){ #map function  iterate over two arguments at a time, norm mdl to have a value and not a list
    if(!is.null(u)){
      p <- predict(u, newdata = data.frame(order = t$order)) #we predict the new values based on the qc normalisation
      t$norm_indiv <- t$log - p + (u$fitted.values[1] + u$residuals[1]) #normalise predicted values by individual slope, trasformed data - predicted value times fitted values plus residuals
      t$norm_indiv_i <- myimputer(v = t$norm_indiv, w = t$value) # impute normalise data
      t
    } else {
      t
    }
  })) 

```


# Feature quality

```{r}
data <- mydat %>%
select(c("name", "data")) %>%
  unnest(data) %>%
  filter(class == "sample") %>% #AGAIN BIG TABLE
select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>% #TAKE NAME from column name and value from norm_indiv_i
  column_to_rownames("id")
mydat$mean_samples <- apply(data, 2, mean)

data <- mydat %>%
  select(c("name", "data")) %>%
  unnest(data) %>%
  filter(class == "blank") %>% #AGAIN BIG TABLE
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>% #TAKE NAME from column name and value from norm_indiv_i
  column_to_rownames("id")
mydat$mean_blank <- apply(data, 2, mean)
mydat$mean_ratio <- mydat$mean_samples / mydat$mean_blank #

rsd <- function(x){(sd(x) / mean(x))*100} #For mass features rsd higher in the qc than in the samples excluded them
data <- mydat %>%
  select(c("name", "data")) %>%
  unnest(data) %>%
  filter(class == "QC") %>% #AGAIN BIG TABLE
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>% #TAKE NAME from column name and value from norm_indiv_i
  column_to_rownames("id")
mydat$RSD_QC <- apply(data, 2, rsd)
data <- mydat %>%
  select(c("name", "data")) %>%
  unnest(data) %>%
  filter(class == "sample") %>% #AGAIN BIG TABLE
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>% #TAKE NAME from column name and value from norm_indiv_i
  column_to_rownames("id")
mydat$RSD_samples <- apply(data, 2, rsd)

mydat$noise <- (mydat$mean_ratio < log10(5)) | (mydat$RSD_QC > mydat$RSD_samples) | 
  (mydat$QC_nonNA < max(mydat$QC_nonNA)*0.75) #Noise if the mean ratio is bellowed 5 or rsd is higher in qc than in samples or feature is zero in sample and blank, is not present in more than 75% of QC

table(mydat$noise)
#rm(data) 
```


# Annotation
#implemente the annotated compounds into the code, run the iondb

## analytical standards annotated features
```{r}
load("rdata/ionsdb_stds.RData")
if(p == "POS"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M+H]+",]
} else if(p == "NEG"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M-H]-",]
}
feat <- data.frame(mydat[,c("name", "mz", "RT", "noise")])
#feat$RT <- as.numeric(feat$RT)
pr <- MzRtParam(tolerance = 0.01, toleranceRt = 0.15) # define the parameters
mt_ft_std <- matchValues(feat, ionsdb, param = pr, 
                     rtColname = c("RT", "ion_rt"),
                     mzColname = c("mz", "ion_mz"))
mtft_df_std <- data.frame(matchedData(mt_ft_std))
tmp_ft <- unique(mtft_df_std$name[which(duplicated(mtft_df_std$name))])
if(length(tmp_ft)>0){
  for(i in seq(length(tmp_ft))){
  idx2 <- which(mtft_df_std$name == tmp_ft[i],)
  tmp <- mtft_df_std[idx2,]
  mtft_df_std <- mtft_df_std[-idx2,]
  if(length(unique(gsub("\\ \\(.*", "", tmp$target_name))) == 1){
    tmp <- tmp[which.min(abs(tmp$score_rt)),]
  } else {
    tmp$target_name <- paste(tmp$target_name, collapse = "; ")
    tmp$target_compound_id <- tmp$target_ion_adduct <- tmp$target_ion_mz <- tmp$target_ion_rt <- tmp$target_comment <- tmp$target_activity <- tmp$target_level <- NA
    tmp <- tmp[1,]
  }
  mtft_df_std <- rbind(mtft_df_std, tmp)
}
}
idx <- which(colnames(mtft_df_std) %in% c("mz", "RT")) #we have to remove the columes with the same name
mtft_df_std <- mtft_df_std[,-idx]

# in case there are duplicates we have to remove one
mtft_df_std <- mtft_df_std %>%
group_by(target_name) %>%
 filter(!(duplicated(target_name) & !is.na(target_name))) %>%  # Remove non-NA duplicates
ungroup()  # Remove the grouping

mydat_std <- mydat %>%
  left_join(mtft_df_std, by = "name")

ft_ms2_std <- mtft_df_std[!is.na(mtft_df_std$target_name),]
write_xlsx(ft_ms2_std, paste0(p, "_analyticalSTDMatch_TableR.xlsx"))

```


## MS2 annotated features

```{r}
load("rdata/ionsdb_AnnComp.RData")
if(p == "POS"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M+H]+",]
} else if(p == "NEG"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M-H]-",]
}
feat <- data.frame(mydat[,c("name", "mz", "RT")])
#feat$RT <- as.numeric(feat$RT)
pr <- MzRtParam(tolerance = 0.035, toleranceRt = 0.15) # define the parameters
mt_ft <- matchValues(feat, ionsdb, param = pr, 
                     rtColname = c("RT", "ion_rt"),
                     mzColname = c("mz", "ion_mz"))
mtft_df <- data.frame(matchedData(mt_ft))
tmp_ft <- unique(mtft_df$name[which(duplicated(mtft_df$name))])
if(length(tmp_ft)>0){
  for(i in seq(length(tmp_ft))){
  idx2 <- which(mtft_df$name == tmp_ft[i],)
  tmp <- mtft_df[idx2,]
  mtft_df <- mtft_df[-idx2,]
  if(length(unique(gsub("\\ \\(.*", "", tmp$target_name))) == 1){
    tmp <- tmp[which.min(abs(tmp$score_rt)),]
  } else {
    tmp$target_name <- paste(tmp$target_name, collapse = "; ")
    tmp$target_compound_id <- tmp$target_ion_adduct <- tmp$target_ion_mz <- tmp$target_ion_rt <- tmp$target_comment <- tmp$target_activity <- tmp$target_level <- NA
    tmp <- tmp[1,]
  }
  mtft_df <- rbind(mtft_df, tmp)
}
}
idx <- which(colnames(mtft_df) %in% c("mz", "RT")) #we have to remove the columes with the same name
mtft_df <- mtft_df[,-idx]

# in case there are duplicates we have to remove one
mtft_df <- mtft_df %>%
group_by(target_name) %>%
 filter(!(duplicated(target_name) & !is.na(target_name))) %>%  # Remove non-NA duplicates
ungroup()  # Remove the grouping

mydat <- mydat %>%
  left_join(mtft_df, by = "name")

ft_ms2 <- mtft_df[!is.na(mtft_df$target_name),]
write_xlsx(ft_ms2, paste0(p, "_AnnotatedComp_TableR.xlsx"))

save(mydat, file = "output/mydat_pos.RData")
#ft_na <- mtft_df[is.na(mtft_df$target_name),]
```


# statistical analysis

## PCA - All samples

```{r}
data <- mydat %>%
  select(c("name", "data", "noise", "target_name")) %>%
  filter(noise == FALSE) %>% #exclude noise
#  filter(!is.na(target_name)) %>% #in case you don't want to only with annotated mass features than uncomment or actually coppy all the pca to have both as well as for volcano plots
  select(c("name", "data"))  %>%
  unnest(data) %>% #AGAIN BIG TABLE
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>% #TAKE NAME from column name and value from norm_indiv_i
  column_to_rownames("id")

data <- apply(data, 2, scaling.pareto) #apply scaling to all colums

pca <- prcomp(data, center = FALSE, scale. = FALSE)
pca_s <- data.frame(pca$x)
s_class <- mydat$data[[1]]$class
plot_ly(data = pca_s, x = ~PC1, y = ~PC2, color = s_class, colors = "Set1",
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC1 (", sprintf("%.1f", summary(pca)$importance[2,1]*100), "%)")),
         yaxis = list(title = paste0(
           "PC2 (", sprintf("%.1f", summary(pca)$importance[2,2]*100), "%)"))
  )
```


## PCA - Study samples 

### All compounds

```{r}
mydat <- mydat %>%
  mutate(data2 = map(data, function(t){
    t$exclude <- grepl("P_|QC|PPB", t$id) # exclude blanks, pooled sampels and qc
    t <- t %>%
      filter(!t$exclude)
    t
  }))


data <- mydat %>%
  select(c("name", "data2", "noise", "target_name")) %>%
  filter(noise == FALSE) %>%
#  filter(!is.na(target_name)) %>% 
  select(c("name", "data2"))  %>%
  unnest(data2) %>%
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>%
  column_to_rownames("id")

data <- apply(data, 2, scaling.pareto)

pca <- prcomp(data, center = FALSE, scale. = FALSE)
pca_s <- data.frame(pca$x)

s_class <- gsub('[[:digit:]]+', '', rownames(data)) #exlude numbers from the names, we had to do it like that because i made some errors in sample names
s_class <- gsub("_", "", s_class)

plot_ly(data = pca_s, x = ~PC1, y = ~PC2, color = s_class, colors = "Set2", size = 2,
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC1 (", sprintf("%.1f", summary(pca)$importance[2,1]*100), "%)")),
         yaxis = list(title = paste0(
           "PC2 (", sprintf("%.1f", summary(pca)$importance[2,2]*100), "%)"))
  )

plot_ly(data = pca_s, x = ~PC3, y = ~PC4, color = s_class, colors = "Set2", size = 2,
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC3 (", sprintf("%.1f", summary(pca)$importance[2,3]*100), "%)")),
         yaxis = list(title = paste0(
           "PC4 (", sprintf("%.1f", summary(pca)$importance[2,4]*100), "%)"))
  )
```


### AnnCompounds

```{r}
mydat <- mydat %>%
  mutate(data2 = map(data, function(t){
    t$exclude <- grepl("P_|QC|PPB", t$id) # exclude blanks, pooled sampels and qc
    t <- t %>%
      filter(!t$exclude)
    t
  }))


data <- mydat %>%
  select(c("name", "data2", "noise", "target_name")) %>%
  filter(noise == FALSE) %>%
 filter(!is.na(target_name)) %>% 
  select(c("name", "data2"))  %>%
  unnest(data2) %>%
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>%
  column_to_rownames("id")

data <- apply(data, 2, scaling.pareto)

pca <- prcomp(data, center = FALSE, scale. = FALSE)
pca_s <- data.frame(pca$x)

s_class <- gsub('[[:digit:]]+', '', rownames(data)) #exlude numbers from the names, we had to do it like that because i made some errors in sample names
s_class <- gsub("_", "", s_class)

plot_ly(data = pca_s, x = ~PC1, y = ~PC2, color = s_class, colors = "Set2", size = 2,
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC1 (", sprintf("%.1f", summary(pca)$importance[2,1]*100), "%)")),
         yaxis = list(title = paste0(
           "PC2 (", sprintf("%.1f", summary(pca)$importance[2,2]*100), "%)"))
  )

plot_ly(data = pca_s, x = ~PC3, y = ~PC4, color = s_class, colors = "Set2", size = 2,
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC3 (", sprintf("%.1f", summary(pca)$importance[2,3]*100), "%)")),
         yaxis = list(title = paste0(
           "PC4 (", sprintf("%.1f", summary(pca)$importance[2,4]*100), "%)"))
  )

save(mydat, file = "output/mydata_annotComp_pos.RData")
```

# Pairwise comparision AnnoComp

```{r}
mydat_annComp <- mydat %>%
  #filter(noise == FALSE) %>%
 filter(!is.na(mean_samples)) %>% #IF THERE IS NO MEAN EXCLUDE
 filter(!is.na(target_name)) %>% 
  mutate(data2 = map(data2, function(t){
    t$class <- gsub('[[:digit:]]+', '', t$id)
    t$class <- gsub("_", "", t$class)
       t
  }))

# Get maximum mean(group):
mydat_annComp <- mydat_annComp %>%
  select(-data) %>%
  mutate(mean_group = map(data2, function(t){
    t %>% 
             group_by(class) %>% 
      summarise(mean_log  = mean(norm_indiv_i))
  })) %>%
  mutate(mean_max = map_dbl(mean_group, function(t){
    t %>%
    # filter(class %in% c("D", "DO", "PF", "PS")) %>%
      pull(mean_log) %>%
      max(.)
  }))
```

## Relevance (based on t-test + effect size):

```{r}
# Relevance (based on t-test + effect size):
mydat_annComp <- mydat_annComp %>%
  mutate(organic_pval = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "DO"))
    t.test(norm_indiv_i ~ class, data = u)$p.value
  })) %>%
  mutate(organic_cd = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "DO"))
    cohens_d(norm_indiv_i ~ class, data = u)$Cohens_d
  })) %>%
  mutate(variety_pval = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "PF"))
    t.test(norm_indiv_i ~ class, data = u)$p.value
  })) %>%
  mutate(variety_cd = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "PF"))
    cohens_d(norm_indiv_i ~ class, data = u)$Cohens_d
  })) %>% 
  mutate(procss_pval = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("PF", "PS"))
    t.test(norm_indiv_i ~ class, data = u)$p.value
  })) %>%
  mutate(procss_cd = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("PF", "PS"))
    cohens_d(norm_indiv_i ~ class, data = u)$Cohens_d
  }))

# indicate if each feature was marker of organic-variety-process or never relevant 
# according to the "new statistics":
mydat_annComp <- mydat_annComp %>%
  mutate(marker_organic = organic_pval < 0.05 & abs(organic_cd) > 1) %>%
  mutate(marker_variety = variety_pval < 0.05 & abs(variety_cd) > 1) %>%
  mutate(marker_procss = procss_pval < 0.05 & abs(procss_cd) > 1) %>%
  mutate(not_marker = (variety_pval > 0.05 | abs(variety_cd) < 1) &
           (procss_pval > 0.05 | abs(procss_cd) < 1) &
           (organic_pval > 0.05 | abs(organic_cd) < 1))

# get the names of markers and not-markers:
mk_organic <- mydat_annComp %>%
  filter(marker_organic) %>%
  arrange(-mean_max) %>%
  pull(name)

mk_variety <- mydat_annComp %>%
  filter(marker_variety) %>%
  arrange(-mean_max) %>%
  pull(name)

mk_procss <- mydat_annComp %>%
  filter(marker_procss) %>%
  arrange(-mean_max) %>%
  pull(name)

mk_NOT <- mydat_annComp %>%
  filter(not_marker) %>%
  arrange(-mean_max) %>%
  pull(name)

save(mydat_annComp, file = "output/mydat_annComp_stat_MK_pos.RData")
write_xlsx(mydat_annComp, path = paste0("output/", p, "_mydat_annComp_stat_MK.xlsx")) #write excel file with normalized intensity of annotated compounds
```


## Plots
### van diagrams
```{r}
# see if markers were common for >1 criteria or not:
venn.plot <- venn.diagram(
  x = list(organic = mk_organic, variety = mk_variety, process = mk_procss, not = mk_NOT),
  category.names = c("Organic", "Variety", "Process", "Not marker"),
  filename = NULL, # Set to NULL to render the plot in the R graphics device
  output = TRUE
)
grid.newpage()
grid.draw(venn.plot)

```

### Heathmaps
```{r}
# Identify markers that are TRUE only once in conditions (excluding NotMarker)

  # List of all markers (combined from all conditions)
markers <- unique(c(mk_organic, mk_variety, mk_procss, mk_NOT))

# Create a data frame where rows are markers and columns are conditions
heatmap_data <- data.frame(
  #Marker = markers,
  Organic = markers %in% mk_organic,  # TRUE if marker is in mk_organic, otherwise FALSE
  Variety = markers %in% mk_variety,  # TRUE if marker is in mk_variety, otherwise FALSE
  Process = markers %in% mk_procss,   # TRUE if marker is in mk_procss, otherwise FALSE
  `NotMarker` = markers %in% mk_NOT, # TRUE if marker is in mk_NOT, otherwise FALSE
  target_name = mydat_annComp$target_name, 
  target_activity = mydat_annComp$target_activity
)
  
  
unique_markers <- heatmap_data %>%
  filter(`NotMarker` == FALSE) %>%  # Exclude NotMarker
  rowwise() %>% 
  mutate(true_count = sum(c(Organic, Variety, Process))) %>%  # Count TRUE values across conditions
  filter(true_count == 1) %>%  # Keep only markers with exactly one TRUE
  pull(target_name)  # Extract the target_name

heatmap_melted <- heatmap_data %>%
  pivot_longer(cols = Organic:NotMarker, 
               names_to = "Condition", 
               values_to = "Status")

# Create a combined column for Status and activity_level to determine colors
heatmap_melted$Status_Activity <- with(heatmap_melted, 
                                       interaction(Status, target_activity, drop = TRUE))

# Plot 1
heatMap <- ggplot(heatmap_melted, aes(x = Condition, y = target_name, fill = Status_Activity)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c(
    "TRUE.protective" = "lightblue",    # Light blue for TRUE and protective
    "TRUE.toxic" = "tomato",           # Red for TRUE and toxic
    "FALSE.protective" = "white",      # White for FALSE and protective
    "FALSE.toxic" = "white"            # White for FALSE and toxic
  )) + 
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),     # Adjust the angle of the x-axis labels
             axis.title.y = element_text(margin = margin(r = 20)),  # Add more margin to the right of y-axis title
    axis.text.y = element_text(margin = margin(r = 20)),   # Add more margin to the right of y-axis labels
    plot.margin = margin(t = 20, r = 40, b = 20, l = 20)    # Add margins around the plot area
  ) +
  labs(x = "Condition", y = "Marker", fill = "Marker Status / Activity") +
  
  # Add stars to target_name only for the condition where it is unique (TRUE only for that condition)
  geom_text(aes(label = ifelse(Status == TRUE & target_name %in% unique_markers & Condition != "NotMarker", "*", "")),
            color = "red", size = 6, hjust = 0.5, vjust = 0.5)
grid.draw(heatMap)


# Plot 2 Create the heatmap and add stars next to the labels for markers with only one TRUE under valid conditions revirsed heathmap
heatMap <- ggplot(heatmap_melted, aes(x = target_name, y = Condition, fill = Status_Activity)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c(
    "TRUE.protective" = "lightblue",    # Light blue for TRUE and protective
    "TRUE.toxic" = "tomato",           # Red for TRUE and toxic
    "FALSE.protective" = "white",      # White for FALSE and protective
    "FALSE.toxic" = "white"            # White for FALSE and toxic
  )) + 
 theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),     # Adjust the angle of the x-axis labels
             axis.title.y = element_text(margin = margin(r = 20)),  # Add more margin to the right of y-axis title
    axis.text.y = element_text(margin = margin(r = 20)),   # Add more margin to the right of y-axis labels
    plot.margin = margin(t = 20, r = 40, b = 20, l = 20)    # Add margins around the plot area
  ) +
  labs(x = "Condition", y = "Marker", fill = "Marker Status / Activity") +
  
  # Add stars to target_name only for the condition where it is unique (TRUE only for that condition)
  geom_text(aes(label = ifelse(Status == TRUE & target_name %in% unique_markers & Condition != "NotMarker", "*", "")),
            color = "red", size = 6, hjust = 0.5, vjust = 0.5)
grid.draw(heatMap)
```

###Boxplots
```{r}
# Boxplots
data_distrib_annotated <- mydat_annComp %>%
  select(c("name", "data2", "target_name")) %>%
 # filter(noise == FALSE) %>%
  filter(!is.na(target_name)) %>% 
  unnest(data2) %>%
  select(c("target_name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "id", values_from = "norm_indiv_i") %>%
  column_to_rownames("target_name")

df <- as.data.frame(t(data_distrib_annotated)) #transform table
df$groups<-rownames(df) #we add coume groups in which we write name of the compunds
df$groups <- gsub('[[:digit:]]+', '', df$groups) #iz stolpca izbrisemo stevilke
df$groups <- gsub("_", "", df$groups)#iz stolpca izbrisemo _
#naredimo for loop, iteriramo cez vsak stolpec razen zadnjega kjer imamo group in izrisemo graf

write_xlsx(df, paste0(p, "_Norm_intensity_AnnComp.xlsx"))#write excel file with normalized intensity of annotated compounds

#box plots with points
for(i in seq(ncol(df)-1)){ 
  y_min <- min(df[[i]], na.rm = TRUE) - 2 
  y_max <- max(df[[i]], na.rm = TRUE) + 2 # Get min value of column and subtract 1
  #naredimo for zanko, spremenljivka i bo zavzela vrednost med 1 in st vrstic -1, i je stevilka stolpca
  print(ggplot(df, aes(x=groups, df[,i])) + 
          theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
          geom_jitter(aes(color=groups), width=0.2, height=0, size=3, show.legend = FALSE)+
          ylim(c(y_min, y_max)) +
          #      ylim(c(0, 10)) +
          xlab("group")+
          ylab("normalized intensity")+
          ggtitle(paste(colnames(df[i]), ionsdb$ion_adduct)))
}
#box plots 
for(i in seq(ncol(df)-1)){ 
  y_min <- min(df[[i]], na.rm = TRUE) - 2 
  y_max <- max(df[[i]], na.rm = TRUE) + 2 # Get min value of column and subtract 1
  #naredimo for zanko, spremenljivka i bo zavzela vrednost med 1 in st vrstic -1, i je stevilka stolpca
  # Create the boxplot
  boxplot(df[[i]] ~ df$groups,  # Boxplot by group
          main = paste("Plot for", colnames(df)[i]),  # Title based on column name
          ylab = "Normalized Intensity",  # Y-axis label
          xlab = "Group",  # X-axis label
          col = "lightblue",  # Box color
          ylim = c(y_min, y_max)  # Dynamic y-axis limits
  )
}
```

## PDF and word documents
```{r}
## Save boxplots as pdf and word documents
#save as pdf
pdf("output/boxplot_anncomp_pos.pdf", paper = "a4", height = 4*3, width = 2*4)
par(mfrow = c(3, 2)) #put three if pdf
for(i in seq(ncol(df)-1)){ 
  y_min <- min(df[[i]], na.rm = TRUE) - 2 
  y_max <- max(df[[i]], na.rm = TRUE) + 2 # Get min value of column and subtract 1
  #naredimo for zanko, spremenljivka i bo zavzela vrednost med 1 in st vrstic -1, i je stevilka stolpca
  # Create the boxplot
  boxplot(df[[i]] ~ df$groups,  # Boxplot by group
          main = paste("Plot for", colnames(df)[i]),  # Title based on column name
          ylab = "Normalized Intensity",  # Y-axis label
          xlab = "Group",  # X-axis label
          col = "lightblue",  # Box color
          ylim = c(y_min, y_max)  # Dynamic y-axis limits
  )
}
dev.off() #closing pdf

#save as word document
# Define the directory where images will be saved
output_dir <- "output"

# Check if the directory exists, if not, create it
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# Create a new Word document
doc <- read_docx()

# Loop through columns (excluding the last one which is for groups)
for(i in seq(ncol(df) - 1)) {
  # Calculate dynamic y-axis limits
  y_min <- min(df[[i]], na.rm = TRUE) - 2
  y_max <- max(df[[i]], na.rm = TRUE) + 2
  
  # Create a ggplot object for the boxplot
  p <- ggplot(df, aes(x = groups, y = df[[i]])) +
    geom_boxplot(fill = "lightblue") +
    labs(title = paste(colnames(df)[i]), 
         y = "Normalized Intensity", 
         x = "Group") +
    ylim(y_min, y_max) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 7),  # Smaller title text size
      axis.title = element_text(size = 7),  # Smaller axis title text size
      axis.text = element_text(size = 5),   # Smaller axis labels text size
      strip.text = element_text(size = 5)   # Smaller strip labels (if any)
    )
  
  # Save the plot to a temporary file (e.g., PNG) in the output directory
  plot_file <- paste0(output_dir, "/boxplot_", colnames(df)[i], ".png")
  ggsave(plot_file, plot = p, width = 3, height = 2)
  
  # Add the plot image to the Word document
  doc <- doc %>%
    body_add_img(src = plot_file, width = 3, height = 2)  # Insert the plot image
  
  # Optionally, remove the plot image file after adding it to the Word document
  file.remove(plot_file)
}

# Save the Word document
print(doc, target = "output/boxplot_anncomp_pos_small.docx")
```


# Session information

```{r session}
Sys.time()-startpoint
#devtools::session_info()
```