---
title: "Annotated comppounds_statistics"
author: "Ana Kovacic"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: false
    number_sections: false
    toc_float: false
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In this document we first combine all annotated compunds processed ander 04_CompAnn_stats and later perform tha statistics and visualisation on the annotated compunds only instead of on the whole dataset. 
**Result:** 
 - we have annotated 78 compounds, but 5 of them are lost, however they were already a little bit questionable, no drama if we skip them
 - 71 compunds were annotated and also are present in the final table
 - from this 71 compounds, 3 are found as noise, but they will still be included because they are found as noise due to negative ratio, but they seem to be ok 
 **done:**
 - perform pairwaise comparision to define the markers for study condition
 - using van diagram show how many of them are markers
 - heatmap to present graphically results, seperate by activity, group and marker specification 
 
# Libraires

```{r}
library(writexl)
library(readxl)
library(tidyverse)
library(plotly)
library(MetaboAnnotation)
library(effectsize)
library(DT)
library(kableExtra)
library(vegan)
library(devtools)
library(reshape2)
library(VennDiagram)
library(grid)
library(officer) 


scaling.pareto <- function(x){(x - mean(x)) / sqrt(sd(x))}
```

# Import data 
```{r}
# Study data import (annotated compounds combine both polarities):
load("output/mydat_neg.RData")
mydat_neg <- mydat
load("output/mydat_pos.RData")
mydat_pos <- mydat
rm(mydat)

mydat_neg_CompAnn <- mydat_neg %>%
  filter(!is.na(target_name))

mydat_neg_CompAnn_filtered <- mydat_neg_CompAnn %>%
  filter(str_detect(target_mode, "\\-"))

mydat_pos_CompAnn <- mydat_pos %>%
  filter(!is.na(target_name))

mydat_pos_CompAnn_filtered <- mydat_pos_CompAnn %>%
  filter(str_detect(target_mode, "\\+"))

combined_df <- bind_rows(mydat_pos_CompAnn_filtered, mydat_neg_CompAnn_filtered)
combined_all <- bind_rows(mydat_pos_CompAnn, mydat_neg_CompAnn)

## use when checking what is combined and what not
# any(combined_all$target_name %in% ionsdb$name)
# 
# load("rdata/ionsdb_AnnComp.RData")
# missing_in_combined_df <- setdiff(ionsdb$name, combined_df$target_name)
# print(missing_in_combined_df)
# 
# missing_in_combined_all <- setdiff(ionsdb$name, combined_all$target_name)
# print(missing_in_combined_all) #check which compounds have you lost: "4-Aminobenzoic acid" "Aspartame", "p-Coumaric acid", "Caffeic aldehyde", "Tomatine (II)"      still, check way, for some rt bad

#remove no need data
rm(combined_all, mydat_neg, mydat_neg_CompAnn_filtered, mydat_pos, mydat_pos_CompAnn, mydat_pos_CompAnn_filtered, mydat_neg_CompAnn)
#save data
write_xlsx(combined_df, file = "output/AnnComp_data.xlsx")
save(combined_df, file = "output/AnnComp_data.RData")
```

# Statistical analysis
```{r}
# Get maximum mean(group):
combined_df <- combined_df %>%
  mutate(data2 = map(data, function(t){
    t$exclude <- grepl("P_|QC|PPB", t$id) # exclude blanks, pooled samples and qc
    t <- t %>%
      filter(!t$exclude)
    t
  }))

mydat_annComp <- combined_df %>%
  mutate(data2 = map(data2, function(t){
    t$class <- gsub('[[:digit:]]+', '', t$id)
    t$class <- gsub("_", "", t$class)
       t
  }))


# Get maximum mean(group):
mydat_annComp <- mydat_annComp %>%
  mutate(mean_group = map(data2, function(t) {
    t %>%
      group_by(class) %>%
      summarise(mean_log = mean(norm_indiv_i), .groups = 'drop')
  })) %>%
  
  # Get maximum mean for each group
  mutate(mean_max = map_dbl(mean_group, function(t) {
    t %>%
      pull(mean_log) %>%
      max()  # Get the maximum mean
  })) %>%
  mutate(mean_max = map_dbl(mean_group, function(t){
    t %>%
    # filter(class %in% c("D", "DO", "PF", "PS")) %>%
      pull(mean_log) %>%
      max(.)
  }))%>%
  mutate(mean_max_class = map_chr(mean_group, function(t){
    t %>%
      filter(mean_log == max(mean_log)) %>%   # Filter for the class with the max mean_log
      pull(class)  # Return the corresponding class name
  }))  %>%
  # Get the second highest mean for each group
  mutate(mean_second_max = map_dbl(mean_group, function(t) {
    t %>%
      pull(mean_log) %>%
      sort(decreasing = TRUE) %>%  # Sort in descending order
      .[2]  # Select the second value (second highest)
  })) %>%

  # Get the class corresponding to the second highest mean
  mutate(mean_second_max_class = map_chr(mean_group, function(t) {
    second_max <- t %>%
      pull(mean_log) %>%
      sort(decreasing = TRUE) %>%
      .[2]  # Get the second highest mean

    # Return the class associated with the second highest mean
    t %>%
      filter(mean_log == second_max) %>%
      pull(class)
  }))

```

## PCA AnnoComp
```{r}
# PCA with annotated compounds and samples only
data <- mydat_annComp %>%
    select(c("name", "data2"))  %>%
  unnest(data2) %>%
  select(c("name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "name", values_from = "norm_indiv_i") %>%
  column_to_rownames("id")

data <- apply(data, 2, scaling.pareto)

pca <- prcomp(data, center = FALSE, scale. = FALSE)
pca_s <- data.frame(pca$x)

s_class <- gsub('[[:digit:]]+', '', rownames(data)) #exlude numbers from the names, we had to do it like that because i made some errors in sample names
s_class <- gsub("_", "", s_class)

plot_ly(data = pca_s, x = ~PC1, y = ~PC2, color = s_class, colors = "Set2", size = 2,
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC1 (", sprintf("%.1f", summary(pca)$importance[2,1]*100), "%)")),
         yaxis = list(title = paste0(
           "PC2 (", sprintf("%.1f", summary(pca)$importance[2,2]*100), "%)"))
  )

plot_ly(data = pca_s, x = ~PC3, y = ~PC4, color = s_class, colors = "Set2", size = 2,
        text = rownames(data)) %>%
  layout(title = "Score plot",
         xaxis = list(title = paste0(
           "PC3 (", sprintf("%.1f", summary(pca)$importance[2,3]*100), "%)")),
         yaxis = list(title = paste0(
           "PC4 (", sprintf("%.1f", summary(pca)$importance[2,4]*100), "%)"))
  )
```

## Pairwise comparision
```{r}
# Relevance (based on t-test + effect size):
mydat_annComp <- mydat_annComp %>%
  mutate(organic_pval = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "DO"))
    t.test(norm_indiv_i ~ class, data = u)$p.value
  })) %>%
  mutate(organic_cd = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "DO"))
    cohens_d(norm_indiv_i ~ class, data = u)$Cohens_d
  })) %>%
  mutate(variety_pval = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "PF"))
    t.test(norm_indiv_i ~ class, data = u)$p.value
  })) %>%
  mutate(variety_cd = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("D", "PF"))
    cohens_d(norm_indiv_i ~ class, data = u)$Cohens_d
  })) %>% 
  mutate(procss_pval = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("PF", "PS"))
    t.test(norm_indiv_i ~ class, data = u)$p.value
  })) %>%
  mutate(procss_cd = map_dbl(data2, function(t){
    u <- t %>%
      filter(class %in% c("PF", "PS"))
    cohens_d(norm_indiv_i ~ class, data = u)$Cohens_d
  }))

# indicate if each feature was marker of organic-variety-process or never relevant 
# according to the "new statistics":
mydat_annComp <- mydat_annComp %>%
  mutate(marker_organic = organic_pval < 0.05 & abs(organic_cd) > 1) %>%
  mutate(marker_variety = variety_pval < 0.05 & abs(variety_cd) > 1) %>%
  mutate(marker_procss = procss_pval < 0.05 & abs(procss_cd) > 1) %>%
  mutate(not_marker = (variety_pval > 0.05 | abs(variety_cd) < 1) &
           (procss_pval > 0.05 | abs(procss_cd) < 1) &
           (organic_pval > 0.05 | abs(organic_cd) < 1))

# get the names of markers and not-markers:
mk_organic <- mydat_annComp %>%
  filter(marker_organic) %>%
  arrange(-mean_max) %>%
  pull(name)

mk_variety <- mydat_annComp %>%
  filter(marker_variety) %>%
  arrange(-mean_max) %>%
  pull(name)

mk_procss <- mydat_annComp %>%
  filter(marker_procss) %>%
  arrange(-mean_max) %>%
  pull(name)

mk_NOT <- mydat_annComp %>%
  filter(not_marker) %>%
  arrange(-mean_max) %>%
  pull(name)

save(mydat_annComp, file = "output/AnnComp_stat_MK.RData")
save(combined_df, file = "output/Acombined_df.RData")

write_xlsx(mydat_annComp, path = paste0("output/mydat_annComp_stat_MK.xlsx")) #write excel file with normalized intensity of annotated compounds
```

## Plots
### van diagrams
```{r}
# see if markers were common for >1 criteria or not:
png(filename = "venn_diagram.png", width = 1200, height = 800, res = 100)
venn.plot <- venn.diagram(
  x = list(organic = mk_organic, variety = mk_variety, process = mk_procss, not = mk_NOT),
  category.names = c("Organic", "Variety", "Process", "Not marker"),
  filename = NULL, # Set to NULL to render the plot in the R graphics device
  output = TRUE,
    cex = 2,  # Adjust the text size (increase for larger text)
  cat.cex = 3 # Optionally make the font bold
)
grid.newpage()
grid.draw(venn.plot)


# Close the PNG device
dev.off()

# Create the Venn diagram for the article only markers
png(filename = "venn_diagram_markers.png", width = 1200, height = 800, res = 100)
venn.plot <- venn.diagram(
  x = list(organic = mk_organic, variety = mk_variety, process = mk_procss),
  category.names = c("Organic", "Variety", "Process"),
  filename = NULL, # Set to NULL to render the plot in the R graphics device
  output = TRUE,
  cex = 2,  # Adjust the text size (increase for larger text)
    cat.cex = 3 # Optionally make the font bold
)

# Draw the plot
grid.newpage()
grid.draw(venn.plot)

# Close the PNG device
dev.off()
```
### HeathMaps
```{r}
# Load the data
data <- read_excel("output/markers_quantitative.xlsx")
data <- as.data.frame(data)


# Calculate average marker values for each group
marker_columns <- c("organic", "variety", "procss", "notmarker")  # Update if needed
averages <- data %>%
  group_by(target_abc) %>%
  summarise(across(all_of(marker_columns), mean, na.rm = TRUE))

# Order groups by type (toxic at top, protective below)
data <- data %>%
  mutate(target_type = ifelse(grepl("toxic", target_activity, ignore.case = TRUE), "toxic", "protective"))

group_order <- data %>%
  select(target_abc, target_type) %>%
  distinct() %>%
  arrange(desc(target_type))

averages <- averages %>%
  left_join(group_order, by = "target_abc") %>%
  arrange(desc(target_type), target_abc)

# Melt the data for heatmap plotting
melted_averages <- melt(averages, id.vars = c("target_abc", "target_type"), variable.name = "Marker", value.name = "Average")

# Create a heatmap
heatmap_plot <- ggplot(melted_averages, aes(x = Marker, y = reorder(target_abc, desc(target_type)), fill = Average)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(round(Average * 100, 1), "%")), size = 3, color = "black") +  # Add average values as percentages inside tiles
  scale_fill_gradient2(low = "tomato", mid = "white", high = "lightgreen", midpoint = 0, na.value = "grey50") +  # Adjust color scale
  theme_minimal() +
  labs(title = "Heatmap of Average Marker Values by Group",
       x = "Marker",
       y = "Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(rows = vars(target_type), scales = "free_y", space = "free_y")  # Separate toxic and protective groups

# Print the heatmap
print(heatmap_plot)
# Save the plot as a PNG file
ggsave("heatmap_plot.png", plot = heatmap_plot, width = 10, height = 8, dpi = 300)

```

##BoxPlots

```{r}
# Boxplots
load("output/AnnComp_stat_MK.RData")

data_distrib_annotated <- mydat_annComp %>%
  select(c("name", "data2", "target_name")) %>%
 # filter(noise == FALSE) %>%
 # filter(!is.na(target_name)) %>% 
  unnest(data2) %>%
  select(c("target_name", "id", "norm_indiv_i")) %>%
  pivot_wider(names_from = "id", values_from = "norm_indiv_i") %>%
  column_to_rownames("target_name")

df <- as.data.frame(t(data_distrib_annotated)) #transform table
df$groups<-rownames(df) #we add coume groups in which we write name of the compunds
df$groups <- gsub('[[:digit:]]+', '', df$groups) #iz stolpca izbrisemo stevilke
df$groups <- gsub("_", "", df$groups)#iz stolpca izbrisemo _
#naredimo for loop, iteriramo cez vsak stolpec razen zadnjega kjer imamo group in izrisemo graf

write_xlsx(df, paste0("Norm_intensity_AnnComp.xlsx"))#write excel file with normalized intensity of annotated compounds

#box plots with points
for(i in seq(ncol(df)-1)){ 
  y_min <- min(df[[i]], na.rm = TRUE) - 2 
  y_max <- max(df[[i]], na.rm = TRUE) + 2 # Get min value of column and subtract 1
  #naredimo for zanko, spremenljivka i bo zavzela vrednost med 1 in st vrstic -1, i je stevilka stolpca
  print(ggplot(df, aes(x=groups, df[,i])) + 
          theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
          geom_jitter(aes(color=groups), width=0.2, height=0, size=3, show.legend = FALSE)+
          ylim(c(y_min, y_max)) +
          #      ylim(c(0, 10)) +
          xlab("group")+
          ylab("normalized intensity")+
          ggtitle(paste(colnames(df[i]), mydat_annComp$target_ion_adduct)))
}
#box plots 
for(i in seq(ncol(df)-1)){ 
  y_min <- min(df[[i]], na.rm = TRUE) - 2 
  y_max <- max(df[[i]], na.rm = TRUE) + 2 # Get min value of column and subtract 1
  #naredimo for zanko, spremenljivka i bo zavzela vrednost med 1 in st vrstic -1, i je stevilka stolpca
  # Create the boxplot
  boxplot(df[[i]] ~ df$groups,  # Boxplot by group
          main = paste("Plot for", colnames(df)[i]),  # Title based on column name
          ylab = "Normalized Intensity",  # Y-axis label
          xlab = "Group",  # X-axis label
          col = "lightblue",  # Box color
          ylim = c(y_min, y_max)  # Dynamic y-axis limits
  )
}
```

### PDF and word documents
```{r}
## Save boxplots as pdf and word documents
#save as pdf
pdf("output/boxplot_anncomp.pdf", paper = "a4", height = 4*3, width = 2*4)
par(mfrow = c(3, 2)) #put three if pdf
for(i in seq(ncol(df)-1)){ 
  y_min <- min(df[[i]], na.rm = TRUE) - 2 
  y_max <- max(df[[i]], na.rm = TRUE) + 2 # Get min value of column and subtract 1
  #naredimo for zanko, spremenljivka i bo zavzela vrednost med 1 in st vrstic -1, i je stevilka stolpca
  # Create the boxplot
  boxplot(df[[i]] ~ df$groups,  # Boxplot by group
          main = paste("Plot for", colnames(df)[i]),  # Title based on column name
          ylab = "Normalized Intensity",  # Y-axis label
          xlab = "Group",  # X-axis label
          col = "lightblue",  # Box color
          ylim = c(y_min, y_max)  # Dynamic y-axis limits
  )
}
dev.off() #closing pdf

#save as word document
# Define the directory where images will be saved
output_dir <- "output"

# Check if the directory exists, if not, create it
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# Create a new Word document
doc <- read_docx()

# Loop through columns (excluding the last one which is for groups)
for(i in seq(ncol(df) - 1)) {
  # Calculate dynamic y-axis limits
  y_min <- min(df[[i]], na.rm = TRUE) - 2
  y_max <- max(df[[i]], na.rm = TRUE) + 2
  
  # Create a ggplot object for the boxplot
  p <- ggplot(df, aes(x = groups, y = df[[i]])) +
    geom_boxplot(fill = "lightblue") +
    labs(title = paste(colnames(df)[i]), 
         y = "Normalized Intensity", 
         x = "Group") +
    ylim(y_min, y_max) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 7),  # Smaller title text size
      axis.title = element_text(size = 7),  # Smaller axis title text size
      axis.text = element_text(size = 5),   # Smaller axis labels text size
      strip.text = element_text(size = 5)   # Smaller strip labels (if any)
    )
  
  # Save the plot to a temporary file (e.g., PNG) in the output directory
  plot_file <- paste0(output_dir, "/boxplot_", colnames(df)[i], ".png")
  ggsave(plot_file, plot = p, width = 3, height = 2)
  
  # Add the plot image to the Word document
  doc <- doc %>%
    body_add_img(src = plot_file, width = 3, height = 2)  # Insert the plot image
  
  # Optionally, remove the plot image file after adding it to the Word document
  file.remove(plot_file)
}

# Save the Word doc
print(doc, target = "output/boxplot_anncomp.docx")
```


# Session information
```{r session}
Sys.time()-startpoint
devtools::session_info()
```
```{r}

```
