---
title: "massFeatures_fitration_effectSize"
author: "AnaK"
date: "2024-05-22"
output: html_document
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
```


# Libraries

```{r libraries}
p <- "POS"
software <- "mzmine" 
myfile <- "data/Fecal_pos_aligned_2e3.xlsx"

library(tidyverse)
library(effectsize)
library(readxl)
library(MetaboAnnotation)
library(plotly)
library(writexl)
library(RColorBrewer)
library(DT)

myimputer <- function(v){
  set.seed(123)
  if(sum(is.na(v)) == 0) {
    return(v)
  } else {
    napos <- which(is.na(v))
    newval <- runif(length(napos), 0, min(v, na.rm = TRUE))
    out <- v
    out[napos] <- newval
    return(out)
  }
}

scaling.pareto <- function(x){(x - mean(x)) / sqrt(sd(x))}
rsd <- function(x){(sd(x) / mean(x))*100}
```

# Data import

```{r import}
if(software == "mzmine"){
  feat <- read_xlsx(myfile) 
  
  data_raw <- feat[,4:ncol(feat)]
  feat <- data.frame(feat[,c(2, 3)])
  colnames(feat) <- c("mz", "RT")
}

feat$FT <- paste0(
  "FT", formatC(seq(nrow(feat)), width = nchar(nrow(feat)), flag = "0"))
rownames(data_raw) <- paste0(
  "FT", formatC(seq(nrow(feat)), width = nchar(nrow(feat)), flag = "0"))
data_raw <- data.frame(t(data_raw)) #transform data
```

# Meta Data Organisation
```{r}
#prepare the metadata of the sample names
md <- gsub(".mzML Peak area", "", rownames(data_raw))
idx <- grep("_QC_0D", md)
md[idx] <- gsub("_QC_0D", "_QC0D", md[idx])
rm(idx)

md <- data.frame(do.call(rbind, strsplit(md, "_")))
colnames(md) <- c("date", "inj_order", "scan", "acquisition", "polarity", "resolution", "donor", "sample", "time")# Name the columns if needed
md$filename <- gsub(".mzML Peak area", "", rownames(data_raw))
idx <- which(md$sample %in% c("QC", "QC0D", "Solvent", "stdMix", "Buffer")) #seperate samples from samples that are not fermentation time dependent 
md$time[idx] <- 9999 #instead of time contains 9999, also names in the pca 
md$time <- as.numeric(gsub("h", "", md$time)) #remove h from time, so that it is numeric and can be use for plots 
rm(idx)
md  <- md 
md$inj_order <- as.numeric(str_replace(md$inj_order, "^0+" ,""))
```

# Raw data
```{r}
if(all(md$filename == gsub(".mzML.*", "", rownames(data_raw)))){
  
} else {
  print("Warning!! Check order of raws between md and data_raw!!!")
}


dt <- cbind(data_raw, md) %>% #bind the raw data with md
  as_tibble() %>% #work with tidyverse, for htis case is better 
  pivot_longer(colnames(data_raw)) %>% #pivot per colnames in order to combine data per feature with nest by name - each feature 
  nest(data = !c("name"))%>%
  mutate(data = map(data, function(t){ #with mutate function you generate a new colume aplying the function on the old one, maping in data per feature a new colume
    t$val_na <- ifelse(t$value == 0, NA, t$value) #a new colume with replaced 0 with na
    t$value_imp <- myimputer(t$val_na) # a new colume with imputed data
    t$val_imp_log <- log10(t$value_imp) #a new colume with transformed data
    t
  }))

# #save data
# save(dt, file = "rdata/dt.RData")
# 
# #if already saves then load data
 load("rdata/dt.RData")
```

# Normalisation - median 1 (batch normalisation)

```{r}
#filter data, remove the first injection, first relevant injection is the last of the first QC with injection order = 8  
dt_normMedian <- dt %>%
  mutate(data_original = data) %>% # the non filtered data 
  mutate(data = map(data, function(t){ #filtered data
    t %>%
      filter(inj_order >= 8)
  }))
#normalization median of each feature = 1
dt_normMedian <- dt_normMedian %>% 
  mutate(data = map(data, function(t){
    t <- t %>%
      group_by(donor) %>%
      mutate(median = median(value_imp),
             val_norm = value_imp / median) 
  }))
load("rdata/dt_normMedian.RData")
#save data
#save(dt_normMedian, file = "rdata/dt_normMedian.RData")
```

# Annotation
## Annotation stds
```{r}
load("rdata/ionsdb_stds.RData")

if(p == "POS"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M+H]+",]
} else if(p == "NEG"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M-H]-",]
}

pr <- MzRtParam(tolerance = 0.01, toleranceRt = 0.15) # define the parameters
mt_ft <- matchValues(feat, ionsdb, param = pr, 
                     rtColname = c("RT", "ion_rt"),
                     mzColname = c("mz", "ion_mz"))
mtft_df <- data.frame(matchedData(mt_ft))
mtft_df <- mtft_df[!is.na(mtft_df$target_name),]

mtft_df$mz <- round(mtft_df$mz, 4) #less decimals
mtft_df$RT <- round(mtft_df$RT, 2)
DT::datatable(mtft_df)
```

## Annotation metabol pathway
```{r}
load("rdata/ionsdb_Metabol_pathway.RData")

if(p == "POS"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M+H]+",]
} else if(p == "NEG"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M-H]-",]
}

pr <- MzParam(tolerance = 0.01) # define the parameters
mt_ft_MP <- matchValues(feat, ionsdb, param = pr, 
                     mzColname = c("mz", "ion_mz"))
mtft_df_MP <- data.frame(matchedData(mt_ft_MP))
mtft_df_MP <- mtft_df_MP[!is.na(mtft_df_MP$target_name),]
mtft_df_MP$mz <- round(mtft_df_MP$mz, 4) #less decimals
mtft_df_MP$RT <- round(mtft_df_MP$RT, 2)
DT::datatable(mtft_df_MP)
```

## Annotation Suspect list

```{r annotation-sl}
load("rdata/ionsdb_SL.RData")
if(p == "POS"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M+H]+",]
} else if(p == "NEG"){
  ionsdb <- ionsdb[ionsdb$ion_adduct == "[M-H]-",]
}

pr <- MzParam(tolerance = 0.01) # define the parameters
mt_ft_sl <- matchValues(feat, ionsdb, param = pr, 
                     mzColname = c("mz", "ion_mz"))
mtft_df_sl <- data.frame(matchedData(mt_ft_sl))
mtft_df_sl <- mtft_df_sl[!is.na(mtft_df_sl$target_name),]
mtft_df_sl$mz <- round(mtft_df_sl$mz, 4) #less decimals
mtft_df_sl$RT <- round(mtft_df_sl$RT, 2)
DT::datatable(mtft_df_sl)
```

# Pairwise comparision - effect size

## Target compounds - ions db 
```{r message=FALSE}
mtft_df$name <- mtft_df$FT
 
 dt_target <- dt %>%
   select(name, data)
 

 dt_target <- dt_target %>%
  left_join(mtft_df)


dt_target <- dt_target[!is.na(dt_target$target_name),]# dt2 is a selection of target compounds, the code otherwise needs some time

#select data based on the condition you want to check
dt_target <- dt_target %>%
  mutate(blank_tomato = map(data, function(t){ #make a new colume maping within data that will contain statistical data named - by the filtered condition and apply the function t
    t %>%
      filter(sample %in% c("Blank", "Tomato1")) %>% #take only the data of interest - control and treatment
      select(c("sample", "val_imp_log", "donor", "time")) %>% #select columes with data you need
      pivot_wider(names_from = sample, values_from = val_imp_log) %>% #pivot the table with sample name as columes and data as rows per selected feature
      group_by(time) %>% #since samples of the same time are dependent we group them by time, to prepare the data for ef and p value of dependent data - paired!
      summarise(ES = cohens_d(Blank, Tomato1, paired = TRUE)$Cohens_d, #a colume with es per time
                pval = t.test(Blank, Tomato1, paired = TRUE)$p.value)#a colume with p value per time
  }))


dt_target <- dt_target %>%
  mutate(blank_tomato = map(blank_tomato, function(t){
    t %>%
      mutate(
        condition = abs(ES) > 1 & pval, # ps more than abs 1, p value less the 0.05
        consecutive_condition = condition & lag(condition, default = NA) #the both conditions must be fulfill by two consequitive time points, true otherwise false, NA if not consequitive, which for time zero can't be, thus giving NA in case it is true
      ) %>%
      mutate(
        consecutive_condition2 = ifelse(time == 0, condition, consecutive_condition) #if NA, than put true, otherwise it is false for zero anyway 
      )
  }))

#both conditions must be true at least within two consequent time points 
dt_target_conseq <- dt_target %>%
  mutate(blank_tomato_relevant = map_lgl(blank_tomato, function(t){
 any(!is.na(t$consecutive_condition) & t$consecutive_condition)
 }))

#at least one time with both conditions is true, if two much we can say one or at least two consequitive 
dt_target_OneTime <- dt_target %>%
  mutate(blank_tomato_relevant = map_lgl(blank_tomato, function(t){
    any(t$condition == TRUE)  # Only consider TRUE values, at least one row is true
  }))
 
#select relevant mass features
 MassFeatures_rawDara_target_oneTime=dt_target_OneTime[dt_target_OneTime$blank_tomato_relevant==TRUE,]
 MassFeatures_rawDara_target_conseqTime=dt_target_conseq[dt_target_conseq$blank_tomato_relevant==TRUE,]

 #write data
 write_csv(MassFeatures_rawDara_target_conseqTime, paste0(p, "MassFeatures_rawDara_target_conseqTime.csv"))
write_csv(MassFeatures_rawDara_target_oneTime, paste0(p, " MassFeatures_rawDara_target_oneTime.csv"))
save(dt_target, file = "rdata/dt_target_raw.RData")
load("rdata/dt_target_raw.RData")

#good to check if the code is doing correct stuff with some manual calculation
#mean1 <- mean(u$Blank)
#mean2 <- mean(u$Aspartame)
#sddif <- sd(u$Blank - u$Aspartame)
#(mean1 - mean2) / sddif
```

## Normalised data - target features 
```{r}
mtft_df$name <- mtft_df$FT
 
 dt_normMedian_target <-  dt_normMedian %>%
   select(name, data)
 

 dt_normMedian_target <- dt_normMedian_target %>%
  left_join(mtft_df)

dt_normMedian_target <- dt_normMedian_target[!is.na(dt_normMedian_target$target_name),]# dt2 is a selection of target compounds, the code otherwise needs some time

#select data based on the condition you want to check
dt_normMedian_target <- dt_normMedian_target %>%
  mutate(blank_tomato = map(data, function(t){ #make a new colume maping within data that will contain statistical data named - by the filtered condition and apply the function t
    t %>%
      filter(sample %in% c("Blank", "Tomato1")) %>% #take only the data of interest - control and treatment
      select(c("sample", "val_norm", "donor", "time")) %>% #select columes with data you need
      pivot_wider(names_from = sample, values_from = val_norm) %>% #pivot the table with sample name as columes and data as rows per selected feature
      group_by(time) %>% #since samples of the same time are dependent we group them by time, to prepare the data for ef and p value of dependent data - paired!
      summarise(ES = cohens_d(Blank, Tomato1, paired = TRUE)$Cohens_d, #a colume with es per time
                pval = t.test(Blank, Tomato1, paired = TRUE)$p.value)#a colume with p value per time
  }))


dt_normMedian_target <- dt_normMedian_target %>%
  mutate(blank_tomato = map(blank_tomato, function(t){
    t %>%
      mutate(
        condition = abs(ES) > 1 & pval, # ps more than abs 1, p value less the 0.05
        consecutive_condition = condition & lag(condition, default = NA) #the both conditions must be fulfill by two consequitive time points, true otherwise false, NA if not consequitive, which for time zero can't be, thus giving NA in case it is true
      ) %>%
      mutate(
        consecutive_condition2 = ifelse(time == 0, condition, consecutive_condition) #if NA, than put true, otherwise it is false for zero anyway 
      )
  }))

#both conditions must be true at least within two consequent time points 
dt_normMedian_target_conseq <- dt_normMedian_target %>%
  mutate(blank_tomato_relevant = map_lgl(blank_tomato, function(t){
 any(!is.na(t$consecutive_condition) & t$consecutive_condition)
 }))

#at least one time with both conditions is true, if two much we can say one or at least two consequitive 
dt_normMedian_target_OneTime <- dt_normMedian_target %>%
  mutate(blank_tomato_relevant = map_lgl(blank_tomato, function(t){
    any(t$condition == TRUE)  # Only consider TRUE values, at least one row is true
  }))
 
#select relevant mass features
 MassFeatures_normDara_target_oneTime=dt_normMedian_target_OneTime[dt_normMedian_target_OneTime$blank_tomato_relevant==TRUE,]
 MassFeatures_normDara_target_conseqTime=dt_normMedian_target_conseq[dt_normMedian_target_conseq$blank_tomato_relevant==TRUE,]

 #write data
 write_csv(MassFeatures_rawDara_target_conseqTime, paste0(p, "MassFeatures_rawDara_target_conseqTime.csv"))
write_csv(MassFeatures_rawDara_target_oneTime, paste0(p, " MassFeatures_rawDara_target_oneTime.csv"))
save(dt_normMedian_target, file = "rdata/dt_target_norm.RData")
load("rdata/dt_neg_normMedian_target_inulin.RData")
```

### some plots of compunds to check
```{r}
#make some curves with selected data in order to see if the statsitical model is selecting relevant mass features, check them with order softwares if they are reall fetaures

idx <- which(dt_neg_normMedian_all_inulin$name == "FT00409")

dt_neg_normMedian_all_inulin$inulin_tomato[idx]

#dt_normMedian_target data for each conditions, per each donor 
dt_neg_normMedian_all_inulin$data[[idx]] %>%
  filter(sample %in% c("Blank", "Tomato1", "Inulin")) %>%
  arrange(donor, sample, time) %>%
 ggplot(aes(x = time, y = val_norm, group = donor, col = donor)) +
  geom_point() + geom_line() + scale_y_log10() + scale_x_continuous(breaks = c(0, 3,6,9,12,24)) +
  facet_wrap(~sample) +
  theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
  xlab("time [h]")+
    ylab("intensity")+
 ggtitle(paste(dt_neg_normMedian_all_inulin$name[idx], "\n", round(dt_neg_normMedian_all_inulin$mz[idx], 6), "@", round(dt_neg_normMedian_all_inulin$RT[idx], 2))) 

#ploting data per each condition all donors, maybe easier to see if there is a time dofference or not 
 dt_normMedian_target$data[[idx]] %>%
  filter(sample %in% c("Blank", "Tomato1","Inulin")) %>%
  arrange(donor, sample, time) %>%
 ggplot(aes(x = time, y = value_imp, group = paste(donor, sample), col = sample)) +
  geom_point() + geom_line() + scale_y_log10() + scale_x_continuous(breaks = c(0, 3,6,9,12,24)) +
theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
  xlab("time [h]")+
    ylab("intensity")+
  ggtitle(paste(dt_normMedian_target$name[idx], "\n", dt_normMedian_target$mz[idx], "@", dt_normMedian_target$RT[idx])) 
```


## All - raw data
```{r message=FALSE}
# mtft_df_sl$name <- mtft_df_sl$FT
#  
#  dt_all <-  dt %>%
#    select(name, data)
#  
# 
#  dt_all <- dt_all %>%
#   left_join(mtft_df_sl)
# 
# #matching with suspect list
# dt_normMedian_all <- dt_normMedian_all %>%
#   left_join(mtft_df_sl %>% select(name, target_name, ppm_error), by = 'name')

dt_all <- dt %>%
  mutate(blank_tomato = map(data, function(t){ #make a new colume maping within data that will contain statistical data named - by the filtered condition and apply the function t
    t %>%
      filter(sample %in% c("Blank", "Tomato1")) %>% #take only the data of interest - control and treatment
      select(c("sample", "val_imp_log", "donor", "time")) %>% #select columes with data you need
      pivot_wider(names_from = sample, values_from = val_imp_log) %>% #pivot the table with sample name as columes and data as rows per selected feature
      group_by(time) %>% #since samples of the same time are dependent we group them by time, to prepare the data for ef and p value of dependent data - paired!
      summarise(ES = cohens_d(Blank, Tomato1, paired = TRUE)$Cohens_d, #a colume with es per time
                pval = t.test(Blank, Tomato1, paired = TRUE)$p.value)#a colume with p value per time
  }))


dt_all <- dt_all %>%
  mutate(blank_tomato = map(blank_tomato, function(t){
    t %>%
      mutate(
        condition = abs(ES) > 1 & pval < 0.05, # ps more than abs 1, p value less the 0.05
        consecutive_condition = condition & lag(condition, default = NA) #the both conditions must be fulfill by two consequitive time points, true otherwise false, NA if not consequitive, which for time zero can't be, thus giving NA in case it is true
      ) %>%
      mutate(
        consecutive_condition2 = ifelse(time == 0, condition, consecutive_condition) #if NA, than put true, otherwise it is false for zero anyway 
      )
  }))


#at least one time with both conditions is true, if two much we can say one or at least two consequitive 
dt_all_oneTrue <- dt_all %>%
  mutate(blank_tomato_relevant = map_lgl(blank_tomato, function(t){
    any(t$condition == TRUE)  # Only consider TRUE values, at least one row is true
  }))

#both conditions must be true at least within two consequent time points 
dt_all_conseq <- dt_all %>%
  mutate(blank_tomato_relevant = map_lgl(blank_tomato, function(t){
 any(!is.na(t$consecutive_condition) & t$consecutive_condition)
 }))

#select relevant data and save
 MassFeatures_rawData_all_conseqTime=dt_all_conseq[dt_all_conseq$blank_tomato_relevant==TRUE,]
 MassFeatures_rawData_target_oneTime=dt_all_oneTrue[dt_all_oneTrue$blank_tomato_relevant==TRUE,]

write_csv( MassFeatures_rawData_all_conseqTime, paste0(p, "_MassFeatures_rawData_all_conseqTime.csv"))
write_csv(MassFeatures_rawData_target_oneTime, paste0(p, "_MassFeatures_rawData_all_oneTime.csv"))
save(dt_all, file = "rdata/dt_all_raw.RData")
```


## Normalised data - all - Median Batch 
```{r message=FALSE}

#select data based on the condition you want to check
dt_normMedian_all_inulin <- dt_normMedian %>%
  mutate(inulin_tomato = map(data, function(t){ #make a new colume maping within data that will contain statistical data named - by the filtered condition and apply the function t
    t %>%
      filter(sample %in% c("Inulin", "Tomato1")) %>% #take only the data of interest - control and treatment
      select(c("sample", "val_norm", "donor", "time")) %>% #select columes with data you need
      pivot_wider(names_from = sample, values_from = val_norm) %>% #pivot the table with sample name as columes and data as rows per selected feature
      group_by(time) %>% #since samples of the same time are dependent we group them by time, to prepare the data for ef and p value of dependent data - paired!
      summarise(ES = cohens_d(Inulin, Tomato1, paired = TRUE)$Cohens_d, #a colume with es per time
                pval = t.test(Inulin, Tomato1, paired = TRUE)$p.value)#a colume with p value per time
  }))

dt_normMedian_all_inulin <- dt_normMedian_all_inulin %>%
  mutate(inulin_tomato = map(inulin_tomato, function(t){
    t %>%
      mutate(
        condition = abs(ES) > 1 & pval, # ps more than abs 1, p value less the 0.05
        consecutive_condition = condition & lag(condition, default = NA) #the both conditions must be fulfill by two consequitive time points, true otherwise false, NA if not consequitive, which for time zero can't be, thus giving NA in case it is true
      ) %>%
      mutate(
        consecutive_condition2 = ifelse(time == 0, condition, consecutive_condition) #if NA, than put true, otherwise it is false for zero anyway 
      )
  }))
save(dt_normMedian_all_inulin, file = "rdata/dt_normMedian_all_inulin.RData")

#both conditions must be true at least within two consequent time points 
dt_normMedian_all_inulin_conseq <- dt_normMedian_all_inulin %>%
  mutate(inulin_tomato_relevant = map_lgl(inulin_tomato, function(t){
 any(!is.na(t$consecutive_condition) & t$consecutive_condition)
 }))
#both conditions must be true for all time points
dt_normMedian_all_inulin_conseqALL <-  dt_normMedian_all_inulin %>%
  mutate(inulin_tomato_relevant = map_lgl(inulin_tomato, function(t){
    all(t$consecutive_condition2)
  }))

#at least one time with both conditions is true, if two much we can say one or at least two consequitive 
dt_normMedian_all_inulin_OneTime <- dt_normMedian_all_inulin %>%
  mutate(inulin_tomato_relevant = map_lgl(inulin_tomato, function(t){
    any(t$condition == TRUE)  # Only consider TRUE values, at least one row is true
  }))
 
#select relevant mass features
 MassFeatures_normData_all_inluin_oneTime=dt_normMedian_all_inulin_OneTime[dt_normMedian_all_inulin_OneTime$inulin_tomato_relevant==TRUE,]
 MassFeatures_normData_all_inluin_conseqTime=dt_normMedian_all_inulin_conseq[dt_normMedian_all_inulin_conseq$inulin_tomato_relevant==TRUE,]
 MassFeatures_normData_all_inluin_ALLTime=dt_normMedian_all_inulin_conseqALL[dt_normMedian_all_inulin_conseqALL$inulin_tomato_relevant==TRUE,]

 #write data
 write_csv(MassFeatures_normData_all_inluin_conseqTime, paste0(p, "_RelevantMassFeatures_normData_tomato_inulin_2conseqTime.csv"))
write_csv(MassFeatures_normData_all_inluin_ALLTime, paste0(p, " _RelevantMassFeatures_normData_tomato_inulin_ALLconseqTime.csv"))
write_csv(MassFeatures_normData_all_inluin_oneTime, paste0(p, " _RelevantMassFeatures_normData_tomato_inulin_oneTime.csv"))

save(dt_normMedian_all, file = "rdata/dt_all_norm.RData")
load("rdata/dt_neg_normMedian_all_inulin.RData")
```

### some plots of kinetics 
```{r}
#make some curves with selected data in order to see if the statsitical model is selecting relevant mass features, check them with order softwares if they are reall fetaures

idx <- which(dt_normMedian_all_inulin$name == "FT00060")

dt_normMedian_all_inulin$inulin_tomato[idx]

#ploted data for each conditions, per each donor 
dt_normMedian_all_inulin$data[[idx]] %>%
  filter(sample %in% c("Inulin", "Tomato1")) %>%
  arrange(donor, sample, time) %>%
 ggplot(aes(x = time, y = value_imp, group = donor, col = donor)) +
  geom_point() + geom_line() + scale_y_log10() + scale_x_continuous(breaks = c(0, 3,6,9,12,24)) +
  facet_wrap(~sample) +
  theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
  xlab("time [h]")+
    ylab("intensity")+
 ggtitle(paste(dt_normMedian_all_inulin$name[idx], "\n", round(dt_normMedian_all_inulin$mz[idx], 6), "@", round(dt_normMedian_all_inulin$RT[idx], 2))) 

#ploting data per each condition all donors, maybe easier to see if there is a time dofference or not 
 dt_normMedian_all_inulin$data[[idx]] %>%
  filter(sample %in% c("Inulin", "Tomato1")) %>%
  arrange(donor, sample, time) %>%
 ggplot(aes(x = time, y = value_imp, group = paste(donor, sample), col = sample)) +
  geom_point() + geom_line() + scale_y_log10() + scale_x_continuous(breaks = c(0, 3,6,9,12,24)) +
theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
  xlab("time [h]")+
    ylab("intensity")+
  ggtitle(paste(dt_normMedian_all_inulin$name[idx], "\n", dt_normMedian_all_inulin$mz[idx], "@", dt_normMedian_all_inulin$RT[idx])) 
 
 idx <- which(dt_normMedian_all_inulin$name == "FT08247")
dt_normMedian_all_inulin$inulin_tomato[idx]
#ploted data for each conditions, per each donor 
dt_normMedian_all$data[[idx]] %>%
  filter(sample %in% c("Inulin", "Tomato1")) %>%
  arrange(donor, sample, time) %>%
 ggplot(aes(x = time, y = value_imp, group = donor, col = donor)) +
  geom_point() + geom_line() + scale_y_log10() + scale_x_continuous(breaks = c(0, 3,6,9,12,24)) +
  facet_wrap(~sample) +
  theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))+
  xlab("time [h]")+
    ylab("intensity")+
 ggtitle(paste(dt_normMedian_all_inulin$name[idx], "\n", round(dt_normMedian_all_inulin$mz[idx], 6), "@", round(dt_normMedian_all_inulin$RT[idx], 2))) 
```

# Session info

```{r session-info}
Sys.time()-startpoint
devtools::session_info()
```